---
title: "XC Analysis"
author: "Samuel Ivanecky"
date: "1/14/2020"
output: html_document
---
# SETUP
```{r setup, include=FALSE}
# Libraries
library(ggplot2)
library(tidyverse)
library(data.table)
library(rvest)
library(forecast)
library(XML)

# Set WD
setwd("~/Desktop/XC/Running/")
```

# GRAB XC RESULTS
```{r}
# To run for D1, womens tables are first except 2013.
# To run for D2, need to flip order of mens and womens tables.
# XC results - 2019
xc19 = 'https://www.tfrrs.org/results/xc/16731/NCAA_Division_I_Cross_Country_Championships' # D1
#xc19 = 'https://www.tfrrs.org/results/xc/16713/NCAA_Division_II_Cross_Country_Championships' # D2
# Read webpage into tables
xc19 = read_html(xc19) %>% html_table()

# Extract each table
women_19 = as.data.frame(xc19[[2]])
men_19 = as.data.frame(xc19[[4]])
# XC results - 2018
xc18 = 'https://www.tfrrs.org/results/xc/15036/NCAA_DI_Cross_Country_Championships' # D1
#xc18 = 'https://www.tfrrs.org/results/xc/15037/NCAA_Division_II_Cross_Country_Championships' # D2

# Read webpage into tables
xc18 = read_html(xc18) %>% html_table()

# Extract each table
women_18 = as.data.frame(xc18[[2]])
men_18 = as.data.frame(xc18[[4]])

# XC results - 2017
xc17 = 'https://www.tfrrs.org/results/xc/13423/NCAA_Division_I_Cross_Country_Championships' # D1
#xc17 = 'https://www.tfrrs.org/results/xc/13406/NCAA_Division_II_Cross_Country_Championships' # D2

# Read webpage into tables
xc17 = read_html(xc17) %>% html_table()

# Extract each table
women_17 = as.data.frame(xc17[[2]])
men_17 = as.data.frame(xc17[[4]])

# XC results - 2016
xc16 = 'https://www.tfrrs.org/results/xc/11271/NCAA_Division_I_Cross_Country_Championships' # D1
#xc16 = 'https://www.tfrrs.org/results/xc/11246/NCAA_Division_II_Cross_Country_Championships' # D2

# Read webpage into tables
xc16 = read_html(xc16) %>% html_table()

# Extract each table
women_16 = as.data.frame(xc16[[2]])
men_16 = as.data.frame(xc16[[4]])

# XC results - 2015
xc15 = 'https://www.tfrrs.org/results/xc/9347/NCAA_Division_I_Cross_Country_Championships' # D1
#xc15 = 'https://www.tfrrs.org/results/xc/9348/NCAA_Division_II_Cross_Country_Championships' # D2

# Read webpage into tables
xc15 = read_html(xc15) %>% html_table()

# Extract each table
women_15 = as.data.frame(xc15[[2]])
men_15 = as.data.frame(xc15[[4]])

# XC results - 2014
xc14 = 'https://www.tfrrs.org/results/xc/7652/NCAA_Division_I_Cross_Country_Championships' # D1
#xc14 = 'https://www.tfrrs.org/results/xc/7658/NCAA_Division_II_National_Championships' # D2

# Read webpage into tables
xc14 = read_html(xc14) %>% html_table()

# Extract each table
women_14 = as.data.frame(xc14[[2]])
men_14 = as.data.frame(xc14[[4]])

# XC results - 2013
xc13 = 'https://www.tfrrs.org/results/xc/6218/NCAA_Division_I_Cross_Country_Championships' # D1
#xc13 = 'https://www.tfrrs.org/results/xc/6221/NCAA_Division_II_Cross_Country_Championships' # D2

# Read webpage into tables
xc13 = read_html(xc13) %>% html_table()

# Extract each table - note order is flipped for D1
# For D2, women are first.
women_13 = as.data.frame(xc13[[4]])
men_13 = as.data.frame(xc13[[2]])

# XC results - 2012
xc12 = 'https://www.tfrrs.org/results/xc/4797/2012_NCAA_Division_I_Cross_Country_Championships' # D1
#xc12 = 'https://www.tfrrs.org/results/xc/4796/2012_NCAA_Division_II_Cross_Country_Championships' # D2

# Read webpage into tables
xc12 = read_html(xc12) %>% html_table()

# Extract each table
women_12 = as.data.frame(xc12[[2]])
men_12 = as.data.frame(xc12[[4]])

# Remove HTML objects
rm(xc12, xc13, xc14, xc15, xc16, xc17, xc18, xc19)

```

# CLEAN UP THE DATA
```{r}
# Only grab necessary columns from each year + add var for finish year
# 2019
men_19 = men_19 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2019)
women_19 = women_19 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2019)
# 2018
men_18 = men_18 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2018)
women_18 = women_18 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2018)
# 2017
men_17 = men_17 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2017)
women_17 = women_17 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2017)
# 2016
men_16 = men_16 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2016)
women_16 = women_16 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2016)
# 2015
men_15 = men_15 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2015)
women_15= women_15 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2015)
# 2014
men_14 = men_14 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2014)
women_14 = women_14 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2014)
# 2013
men_13 = men_13 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2013)
women_13 = women_13 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2013)
# 2012
men_12 = men_12 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2012)
women_12 = women_12 %>%
  select(NAME, TEAM, YEAR, TIME, PL) %>%
  mutate(MEET_YR = 2012)
```

# CREATE OVERALL DATA SET
```{r}
men = rbind(men_12, men_13, men_14, men_15, men_16, men_17, men_18, men_19)
women = rbind(women_12, women_13, women_14, women_15, women_16, women_17, women_18, women_19)
```

# ONLY FOR D2
# CODE TO ADJUST NAMES FOR PROPER GROUPING
```{r}
# men = separate(men, NAME, c("LAST", "FIRST"))
# men = men %>%
#   mutate(
#     NAME = paste0(FIRST, " ", LAST)
#   ) %>%
#   select(NAME, TEAM, YEAR, PL)
# 
# women = separate(women, NAME, c("LAST", "FIRST"))
# women = women %>%
#   mutate(
#     NAME = paste0(FIRST, " ", LAST)
#   ) %>%
#   select(NAME, TEAM, YEAR, PL)
```

# CREATE YEAR PLACE VARS
```{r}
# Men
men$PL = as.numeric(men$PL)
men_grp = men %>%
  mutate(
    FR.PL = case_when(
      grepl("FR", YEAR) ~ PL,
      T ~ 999
    ),
    SO.PL = case_when(
      grepl("SO", YEAR) ~ PL,
      T ~ 999
    ),
    JR.PL = case_when(
      grepl("JR|3", YEAR) ~ PL,
      T ~ 999
    ),
   SR.PL = case_when(
     grepl("SR", YEAR) ~ PL,
     T ~ 999
   )
  ) %>%
  group_by(NAME) %>%
  summarise(TEAM = min(TEAM), FR = min(FR.PL), SO = min(SO.PL), JR = min(JR.PL), SR = min(SR.PL))

# Women
women$PL = as.numeric(women$PL)
women_grp = women %>%
  mutate(
    FR.PL = case_when(
      grepl("FR", YEAR) ~ PL,
      T ~ 999
    ),
    SO.PL = case_when(
      grepl("SO", YEAR) ~ PL,
      T ~ 999
    ),
    JR.PL = case_when(
      grepl("JR|3", YEAR) ~ PL,
      T ~ 999
    ),
   SR.PL = case_when(
     grepl("SR", YEAR) ~ PL,
     T ~ 999
   )
  ) %>%
  group_by(NAME) %>%
  summarise(TEAM = max(TEAM), FR = min(FR.PL), SO = min(SO.PL), JR = min(JR.PL), SR = min(SR.PL))
```

# HANDLE -999
```{r}
men_grp$FR = ifelse(men_grp$FR == 999, NA, men_grp$FR)
men_grp$SO = ifelse(men_grp$SO == 999, NA, men_grp$SO)
men_grp$JR = ifelse(men_grp$JR == 999, NA, men_grp$JR)
men_grp$SR = ifelse(men_grp$SR == 999, NA, men_grp$SR)

women_grp$FR = ifelse(women_grp$FR == 999, NA, women_grp$FR)
women_grp$SO = ifelse(women_grp$SO == 999, NA, women_grp$SO)
women_grp$JR = ifelse(women_grp$JR == 999, NA, women_grp$JR)
women_grp$SR = ifelse(women_grp$SR == 999, NA, women_grp$SR)
```

# WRITE TO CSV
```{r}
write.csv(men_grp, "MEN_XC_All_Time_D1.csv")
write.csv(women_grp, "WOMEN_XC_All_Time_D1.csv")
```

# EDA
```{r}
# MEN
ggplot(men_grp, aes(FR)) + geom_histogram(bins = 25, colour = "blue", fill = "lightblue") + ggtitle("Distribution of Male FR Finishes") + labs(x = "Finish", y = "Count")
ggplot(men_grp, aes(SO)) + geom_histogram(bins = 25, colour = "blue", fill = "lightblue") + ggtitle("Distribution of Male SO Finishes") + labs(x = "Finish", y = "Count")
ggplot(men_grp, aes(JR)) + geom_histogram(bins = 25, colour = "blue", fill = "lightblue") + ggtitle("Distribution of Male JR Finishes") + labs(x = "Finish", y = "Count")
ggplot(men_grp, aes(SR)) + geom_histogram(bins = 25, colour = "blue", fill = "lightblue") + ggtitle("Distribution of Male SR Finishes") + labs(x = "Finish", y = "Count")

# WOMEN
ggplot(women_grp, aes(FR)) + geom_histogram(bins = 25, colour = "blue", fill = "lightblue") + ggtitle("Distribution of Female FR Finishes") + labs(x = "Finish", y = "Count")
ggplot(women_grp, aes(SO)) + geom_histogram(bins = 25, colour = "blue", fill = "lightblue") + ggtitle("Distribution of Female SO Finishes") + labs(x = "Finish", y = "Count")
ggplot(women_grp, aes(JR)) + geom_histogram(bins = 25, colour = "blue", fill = "lightblue") + ggtitle("Distribution of Female JR Finishes") + labs(x = "Finish", y = "Count")
ggplot(women_grp, aes(SR)) + geom_histogram(bins = 25, colour = "blue", fill = "lightblue") + ggtitle("Distribution of Female SR Finishes") + labs(x = "Finish", y = "Count")

```

# CREATE AN ALL TIME RESULTS DF
```{r}
men$GENDER = "M"
women$GENDER = "F"
results = rbind(men, women)
```

# FEATURE ENGINEERING
## CREATE GENDER VAR
```{r}
# Gender var
men_grp$GENDER = "M"
women_grp$GENDER = "F"

# Join two tables together
overall = rbind(men_grp, women_grp)
```

## GET JUMPS PER YEAR
```{r}
overall = overall %>%
  mutate(
    J1 = FR-SO,
    J2 = SO-JR,
    J3 = JR-SR
  )
```

## DISTRIBUTION OF JUMPS
```{r}
ggplot(overall, aes(J1)) + geom_histogram(bins = 35, fill = "lightblue", color = "blue") + ggtitle("Distribution of Jump Between FR & SO YRs")
ggplot(overall, aes(J2)) + geom_histogram(bins = 35, fill = "lightblue", color = "blue") + ggtitle("Distribution of Jump Between SO & JR YRs")
ggplot(overall, aes(J3)) + geom_histogram(bins = 35, fill = "lightblue", color = "blue") + ggtitle("Distribution of Jump Between JR & SR YRs")
```

## DISTRIBUTION OF FINISH BY GRADE
```{r}
# Freshman
fr = ggplot(overall) +
  geom_density(alpha = 0.4, aes(FR, fill = GENDER)) + 
  ggtitle("Distribution of Freshman Finishes") + 
  labs(x = "Finish", y = "Density")

# Sophomore
so = ggplot(overall) +
  geom_density(alpha = 0.4, aes(SO, fill = GENDER)) + 
  ggtitle("Distribution of Sophomore Finishes") + 
  labs(x = "Finish", y = "Density")

# Junior
jr = ggplot(overall) +
  geom_density(alpha = 0.4, aes(JR, fill = GENDER)) + 
  ggtitle("Distribution of Junior Finishes") + 
  labs(x = "Finish", y = "Density")

# Senior
sr = ggplot(overall) +
  geom_density(alpha = 0.4, aes(SR, fill = GENDER)) + 
  ggtitle("Distribution of Senior Finishes") + 
  labs(x = "Finish", y = "Density")

gridExtra::grid.arrange(fr, so, jr, sr)
```

## DISTRIBUTION OF JUMPS BY GENDER
```{r}
# Jump 1
j1 = ggplot(overall) +
  geom_density(alpha = 0.4, aes(J1, fill = GENDER)) + 
  ggtitle("Jump from FR to SO") + 
  labs(x = "Jump in Finish", y = "Density")

# Jump 2
j2 = ggplot(overall) +
  geom_density(alpha = 0.4, aes(J2, fill = GENDER)) + 
  ggtitle("Jump from SO to JR") + 
  labs(x = "Jump in Finish", y = "Density")

# Jump 3
j3 = ggplot(overall) +
  geom_density(alpha = 0.4, aes(J3, fill = GENDER)) + 
  ggtitle("Jump from JR to SR") + 
  labs(x = "Jump in Finish", y = "Density")

gridExtra::grid.arrange(j1, j2, j3)
```

# GROUP ON SCHOOL LEVEL
```{r}
team_overall = overall %>%
  #filter(GENDER == 'F') %>%
  group_by(TEAM) %>%
  summarise(NUM_RUNNERS = n_distinct(NAME), 
            FR.AVG = round(mean(FR, na.rm = T), 2), 
            SO.AVG = round(mean(SO, na.rm = T), 2), 
            JR.AVG = round(mean(JR, na.rm = T), 2), 
            SR.AVG = round(mean(SR, na.rm = T), 2),
            NUM.FR = n_distinct(NAME[!is.na(FR)]), 
            NUM.SO = n_distinct(NAME[!is.na(SO)]),
            NUM.JR = n_distinct(NAME[!is.na(JR)]),
            NUM.SR = n_distinct(NAME[!is.na(SR)]))
```

## SHOW TOP FRESHMAN SCHOOLS
```{r}
# Aggregate by Freshman Finish
freshman = team_overall %>%
  filter(NUM.FR >= 3) %>%
  arrange(FR.AVG)

# Only show top 10
freshman = freshman[1:10, ]

# PLot
ggplot(freshman, aes(reorder(TEAM, FR.AVG), FR.AVG, fill = TEAM)) + 
  geom_bar(stat = "identity") + 
  ggtitle(label = "Best Average Freshman Finish by School", subtitle = "Schools must have at least 3 freshman finishes.") +
  labs(x = "School", y = "Average Finish") +
  geom_text(aes(label=FR.AVG), position=position_dodge(width=0.9), vjust=-0.25) +
  guides(fill=FALSE)

# Aggregate by number of freshman runners
freshman = team_overall %>%
  arrange(-NUM.FR)

# Onlt take top 10
freshman = freshman[1:25, ]

# Plot
ggplot(freshman, aes(reorder(TEAM, -NUM.FR), NUM.FR, fill = TEAM)) + 
  geom_bar(stat = "identity") + 
  ggtitle(label = "Most Freshman Finishes by School") +
  labs(x = "School", y = "Count of Freshman Finishes") +
  geom_text(aes(label=NUM.FR), position=position_dodge(width=0.9), vjust=-0.25) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill=FALSE)
```

# SUBSET TO ONLY RUNNERS WHO HAVE FINISHES FOR ALL 4 YEARS
```{r}
four_years = overall %>%
  filter(!(is.na(FR) | is.na(SO) | is.na(JR) | is.na(SR))) %>%
  #filter(GENDER == "M") %>%
  mutate(
    AVG_FINISH = (FR + SO + JR + SR) / 4
  ) %>%
  arrange(AVG_FINISH)
```

## PLOT OUT TOP RUNNERS BY AVG FINISH
```{r}
# Subset to only top 50
top50 = four_years
# Plot
ggplot(top50, aes(reorder(NAME, AVG_FINISH), AVG_FINISH, fill = NAME)) + 
  geom_bar(stat = "identity") + 
  ggtitle(label = "Best Average Finish", subtitle = "Only runners who have 4 years at NCAAs") +
  labs(x = "Runner", y = "Average Finish") +
  geom_text(aes(label=AVG_FINISH), position=position_dodge(width=0.9), hjust = -0.05, angle = 90) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill=FALSE)
```

```{r}
# Group to team level
team_avg = four_years %>%
  group_by(TEAM) %>%
  summarise(AVG_FINISH = round(mean(AVG_FINISH), 2), NUM_RUNNERS = n_distinct(NAME)) %>%
  filter(NUM_RUNNERS > 1)

# Plot
ggplot(team_avg, aes(reorder(TEAM, AVG_FINISH), AVG_FINISH, fill = TEAM)) + 
  geom_bar(stat = "identity") + 
  ggtitle(label = "Best Average Finish - Team", subtitle = "Only teams with multiple runners") +
  labs(x = "Team", y = "Average Finish") +
  geom_text(aes(label=AVG_FINISH), position=position_dodge(width=0.9), hjust = -0.05, angle = 90) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill=FALSE)

```

# ANALYSIS OF ALL-TIME RESULTS
## DATA CLEANING
```{r}
results$MEET_YR = as.factor(results$MEET_YR)

results = results %>%
  mutate(
    YEAR = case_when(
      grepl("FR|1", YEAR) ~ "FR",
      grepl("SO|2", YEAR) ~ "SO",
      grepl("JR|3", YEAR) ~ "JR",
      grepl("SR|4", YEAR) ~ "SR",
      T ~ "OTHER"
    )
  ) %>%
  mutate(
    TIME = as.numeric(substr(TIME, 1, 2)) * 60 + as.numeric(substr(TIME, 4, 5))
  )
```

## NORMALIZED TIME
```{r}
# Get average for each year by gender
# Men
men.avg2012 = round(mean(results[which(results$MEET_YR == '2012' & results$GENDER == 'M'), ]$TIME, na.rm = T), 2)
men.avg2013 = round(mean(results[which(results$MEET_YR == '2013' & results$GENDER == 'M'), ]$TIME, na.rm = T), 2)
men.avg2014 = round(mean(results[which(results$MEET_YR == '2014' & results$GENDER == 'M'), ]$TIME, na.rm = T), 2)
men.avg2015 = round(mean(results[which(results$MEET_YR == '2015' & results$GENDER == 'M'), ]$TIME, na.rm = T), 2)
men.avg2016 = round(mean(results[which(results$MEET_YR == '2016' & results$GENDER == 'M'), ]$TIME, na.rm = T), 2)
men.avg2017 = round(mean(results[which(results$MEET_YR == '2017' & results$GENDER == 'M'), ]$TIME, na.rm = T), 2)
men.avg2018 = round(mean(results[which(results$MEET_YR == '2018' & results$GENDER == 'M'), ]$TIME, na.rm = T), 2)
men.avg2019 = round(mean(results[which(results$MEET_YR == '2019' & results$GENDER == 'M'), ]$TIME, na.rm = T), 2)

# Women
women.avg2012 = round(mean(results[which(results$MEET_YR == '2012' & results$GENDER == 'F'), ]$TIME, na.rm = T), 2)
women.avg2013 = round(mean(results[which(results$MEET_YR == '2013' & results$GENDER == 'F'), ]$TIME, na.rm = T), 2)
women.avg2014 = round(mean(results[which(results$MEET_YR == '2014' & results$GENDER == 'F'), ]$TIME, na.rm = T), 2)
women.avg2015 = round(mean(results[which(results$MEET_YR == '2015' & results$GENDER == 'F'), ]$TIME, na.rm = T), 2)
women.avg2016 = round(mean(results[which(results$MEET_YR == '2016' & results$GENDER == 'F'), ]$TIME, na.rm = T), 2)
women.avg2017 = round(mean(results[which(results$MEET_YR == '2017' & results$GENDER == 'F'), ]$TIME, na.rm = T), 2)
women.avg2018 = round(mean(results[which(results$MEET_YR == '2018' & results$GENDER == 'F'), ]$TIME, na.rm = T), 2)
women.avg2019 = round(mean(results[which(results$MEET_YR == '2019' & results$GENDER == 'F'), ]$TIME, na.rm = T), 2)

# Create normalized times
results = results %>%
  mutate(
    NORM_TIME = case_when(
      GENDER == 'F' ~ case_when(
        MEET_YR == '2012' ~ TIME / women.avg2012,
        MEET_YR == '2013' ~ TIME / women.avg2013,
        MEET_YR == '2014' ~ TIME / women.avg2014,
        MEET_YR == '2015' ~ TIME / women.avg2015,
        MEET_YR == '2016' ~ TIME / women.avg2016,
        MEET_YR == '2017' ~ TIME / women.avg2017,
        MEET_YR == '2018' ~ TIME / women.avg2018,
        T ~ TIME / women.avg2019
      ),
      GENDER == 'M' ~ case_when(
        MEET_YR == '2012' ~ TIME / men.avg2012,
        MEET_YR == '2013' ~ TIME / men.avg2013,
        MEET_YR == '2014' ~ TIME / men.avg2014,
        MEET_YR == '2015' ~ TIME / men.avg2015,
        MEET_YR == '2016' ~ TIME / men.avg2016,
        MEET_YR == '2017' ~ TIME / men.avg2017,
        MEET_YR == '2018' ~ TIME / men.avg2018,
        T ~ TIME / men.avg2019
      ),
      T ~ 0
    )
  )

```

## DISTRIBUTION OF FINISH BY YEAR BY GRADE
```{r}
ggplot(results, aes(PL, fill = MEET_YR)) +
  geom_histogram(bins = 35) +
  facet_wrap(YEAR ~ .)
```

```{r}
# Womens
ggplot(results[which(results$GENDER == "F"), ], aes(MEET_YR, NORM_TIME, colour = MEET_YR)) +
  geom_boxplot() +
  guides(colour = FALSE) + 
  ylim(NA, 1.1) +
  facet_wrap(YEAR ~ .) +
  ggtitle("Distributions of Female Results from NCAAs by Grade") + 
  labs(x = "Meet", y = "Normalized Time")

ggplot(results[which(results$GENDER == "F"), ], aes(NORM_TIME, fill = YEAR)) +
  geom_density(alpha = 0.4) +
  facet_wrap(MEET_YR ~ .) + 
  xlim(NA, 1.1) +
  ggtitle("Distributions of Results by Grade Over the Years - Female") +
  labs(x = "Normalized Time", y = "Density")

# Mens
ggplot(results[which(results$GENDER == "M"), ], aes(MEET_YR, NORM_TIME, colour = MEET_YR)) +
  geom_boxplot() +
  guides(colour = FALSE) + 
  ylim(NA, 1.1) +
  facet_wrap(YEAR ~ .) +
  ggtitle("Distributions of Male Results from NCAAs by Grade") + 
  labs(x = "Meet", y = "Normalized Time")

ggplot(results[which(results$GENDER == "M"), ], aes(NORM_TIME, fill = YEAR)) +
  geom_density(alpha = 0.4) +
  facet_wrap(MEET_YR ~ .) + 
  ggtitle("Distributions of Results by Grade Over the Years - Male") +
  labs(x = "Normalized Time", y = "Density")
 
```

# DYNAMICALLY SCRAPE TFRRS LINKS
## getURLs
```{r}
# TFRRS Results Page
#url = "https://www.tfrrs.org/results_search.html"
#url = 'https://www.tfrrs.org/results/xc/16731/NCAA_Division_I_Cross_Country_Championships'

getURLs = function(url) {
  # Read in the links on the webpage
  wp = read_html(url) %>%
    html_nodes(xpath = "//td/div/a") %>%
    html_attr("href")
  
  # Manipulate strings
  for ( i  in 1:length(wp) )
  {
    temp = wp[i]
    #temp = substring(temp, 3)
    temp = paste0("https:", temp)
    temp = substr(temp, 1, nchar(temp)-3)
    wp[i] = temp
  }
  
  # Grab only strings for athletes
  wp = wp[grepl("athletes", wp)]
  
  # Return webpage results
  return(wp)

}
```

## runnerScrape Function
```{r}
runnerScrape = function(url){
  # Attempt to read in first runner
  runner = read_html(url) %>%
    html_nodes("tr") %>%
    html_text()
  
  # Get the name of the runner off of TFRRS HTML
  runner_name = read_html(url) %>%
    html_nodes("h3") %>%
    html_text()
  
  runner_name = unlist(strsplit(runner_name[1], "[\n]"))[1]
  
  runner = gsub("[\t\n]", " ", runner)
  
  # Temporary data frame
  athlete = data.frame(cbind("A", "B", "C"))
  names(athlete) = c("EVENT", "TIME", "PLACE")
  
  for ( i in 1:length(runner) )
  {
    # Create temporary string
    temp = unlist(strsplit(runner[i], " "))
    # Check to make sure temp has char
    if(length(temp) > 0)
    {
      # Remove empty cells
      temp = temp[nchar(temp) > 0]
      # Make sure length is greater than 3 and not greater than 4
      if (length(temp) >= 3 & length(temp) <= 4)
      {
        # Take first three eleements
        temp = as.data.frame(cbind(temp[1], temp[2], temp[3]))
        names(temp) = c("EVENT", "TIME", "PLACE")
        # Add to the data frame
        athlete = rbind(athlete, temp)
      }
    }
  }
  
  # Convert to character
  athlete$PLACE = as.character(athlete$PLACE)
  
  # Clean up results
  athlete = athlete %>%
    filter(grepl(":", TIME))
  
  # Clean up and group to one row per event
  athlete = athlete %>%
    mutate(
      EVENT = case_when(
        grepl("5000|5K|5k", EVENT) & !(grepl("XC|xc|Xc", EVENT)) ~ "5000m",
        grepl("3000|3K|3k", EVENT) & !grepl("3000S|3000mS|3000SC", EVENT) ~ "3000m",
        grepl("mile|Mile|MILE", EVENT) ~ "Mile",
        grepl("6K|6k|6000", EVENT) ~ "6K XC",
        grepl("800|800m", EVENT) & !grepl("8000m|8000", EVENT) ~ "800m",
        grepl("8k|8K|8000m", EVENT) & !grepl("\\.", EVENT) ~ "8K XC",
        grepl("10,000|10K|10k|10000", EVENT) & !(grepl("XC|xc|Xc", EVENT)) ~ "10K",
        grepl("10,000|10K|10k|10000", EVENT) & grepl("XC|xc|Xc", EVENT) ~ "10K XC",
        grepl("1500|1500m", EVENT) & !grepl("4x", EVENT) ~ "1500m",
        grepl("3000S|3000s|3000SC|3000sc|3000mS", EVENT) ~ "3000S",
        T ~ "OTHER"
      )
    ) %>%
    mutate(
      PLACE = as.numeric(gsub("th|TH|st|ST|nd|ND|rd|RD", "", PLACE))
    ) %>%
    filter(!is.na(PLACE)) %>%
    filter(PLACE <= 300) %>%
    mutate(TIME = as.character(TIME)) %>%
    mutate(
      TIME = case_when(
       nchar(TIME) == 8 | substr(TIME, 6, 6) == '.' ~ as.numeric(substr(TIME, 1, 2)) * 60 + as.numeric(substr(TIME, 4, 5)),
       nchar(TIME) == 7 ~ as.numeric(substr(TIME, 1, 1)) * 60 + as.numeric(substr(TIME, 3, 4)),
       T ~ 0
      )
    ) %>%
    group_by(EVENT) %>%
    summarise(
      AVG_PLACE = round(mean(PLACE, na.rm = T), 2),
      AVG_TIME = round(mean(TIME), 2),
      PR = min(TIME),
      WINS = n_distinct(TIME[PLACE == 1]),
      TIMES.RUN = n()
    ) %>%
    mutate(
      WIN_PCT = round(((WINS / TIMES.RUN) * 100), 2)
    )
  
  athlete$NAME = runner_name
  
  return(athlete)
}
```

## resultsQuery Function
```{r}
# Create a function for results querying - takes in a list of webpage URLs from runnerScrape
resultsQuery = function(wp){
  # Create a temporary dataframe
  runners = as.data.frame(cbind("event", 1.1, 1.1, 1.1, 1.1, 1.1, 1.1, "name"))
  # Rename columns
  names(runners) = c("EVENT", "AVG_PLACE", "AVG_TIME", "PR", "WINS", "TIMES.RUN", "WIN_PCT", "NAME")
  # Reformat var
  runners = runners %>%
    mutate(
      AVG_PLACE = as.numeric(AVG_PLACE),
      AVG_TIME = as.numeric(AVG_TIME),
      PR = as.numeric(PR),
      WINS = as.numeric(WINS),
      TIMES.RUN = as.numeric(TIMES.RUN),
      WIN_PCT = as.numeric(WIN_PCT),
    )
  
  # Loop through every person from the results
  for ( i in 1:length(wp) )
  {
    # Set url to scrape
    url = wp[i]
    
    # Make a temporary df for the runner
    runner_temp = runnerScrape(url)
    
    # Bind this to the runners df
    runners = rbind(runners, runner_temp)
  }
  
  # Return data frame
  return(runners)
}
```

## reformatRunners
```{r}
# Create function to convert times and transpose
reformatRunners = function(df){
  df = df %>%
    mutate( # Create times that are readable for QC purposes
      AVG_TIME_FORM = paste0(floor(AVG_TIME/60), ":", str_pad(round((AVG_TIME%%60), 2), width = 2, side = "left", pad = "0"))
    ) %>%
    mutate(
      PR_FORM = paste0(floor(PR/60), ":", str_pad(round((PR%%60), 2), width = 2, side = "left", pad = "0"))
    ) %>%
    filter(!(EVENT %in% c("event", "OTHER"))) %>%
    group_by(NAME) %>%
    summarise( # Transpose data frame
      # 800m
      AVG_PLACE_800 = min(AVG_PLACE[EVENT == '800m']),
      AVG_TIME_800 = min(AVG_TIME[EVENT == '800m']),
      PR_800 = min(PR[EVENT == '800m']),
      WINS_800 = min(WINS[EVENT == '800m']),
      TIMES_RUN_800 = min(TIMES.RUN[EVENT == '800m']),
      WIN_PCT_800 = min(WIN_PCT[EVENT == '800m']),
      AVG_TIME_FORM_800 = min(AVG_TIME_FORM[EVENT == '800m']),
      PR_FORM_800 = min(PR_FORM[EVENT == '800m']),
      # 1500m
      AVG_PLACE_1500 = min(AVG_PLACE[EVENT == '1500m']),
      AVG_TIME_1500 = min(AVG_TIME[EVENT == '1500m']),
      PR_1500 = min(PR[EVENT == '1500m']),
      WINS_1500 = min(WINS[EVENT == '1500m']),
      TIMES_RUN_1500 = min(TIMES.RUN[EVENT == '1500m']),
      WIN_PCT_1500 = min(WIN_PCT[EVENT == '1500m']),
      AVG_TIME_FORM_1500 = min(AVG_TIME_FORM[EVENT == '1500m']),
      PR_FORM_1500 = min(PR_FORM[EVENT == '1500m']),
      # Mile
      AVG_PLACE_MILE = min(AVG_PLACE[EVENT == 'Mile']),
      AVG_TIME_MILE = min(AVG_TIME[EVENT == 'Mile']),
      PR_MILE = min(PR[EVENT == 'Mile']),
      WINS_MILE = min(WINS[EVENT == 'Mile']),
      TIMES_RUN_MILE = min(TIMES.RUN[EVENT == 'Mile']),
      WIN_PCT_MILE = min(WIN_PCT[EVENT == 'Mile']),
      AVG_TIME_FORM_MILE = min(AVG_TIME_FORM[EVENT == 'Mile']),
      PR_FORM_MILE = min(PR_FORM[EVENT == 'Mile']),
      # 3000m
      AVG_PLACE_3000 = min(AVG_PLACE[EVENT == '3000m']),
      AVG_TIME_3000 = min(AVG_TIME[EVENT == '3000m']),
      PR_3000 = min(PR[EVENT == '3000m']),
      WINS_3000 = min(WINS[EVENT == '3000m']),
      TIMES_RUN_3000 = min(TIMES.RUN[EVENT == '3000m']),
      WIN_PCT_3000 = min(WIN_PCT[EVENT == '3000m']),
      AVG_TIME_FORM_3000 = min(AVG_TIME_FORM[EVENT == '3000m']),
      PR_FORM_3000 = min(PR_FORM[EVENT == '3000m']),
      # 3000mSC
      AVG_PLACE_3000S = min(AVG_PLACE[EVENT == '3000S']),
      AVG_TIME_3000S = min(AVG_TIME[EVENT == '3000S']),
      PR_3000S = min(PR[EVENT == '3000S']),
      WINS_3000S = min(WINS[EVENT == '3000S']),
      TIMES_RUN_3000S = min(TIMES.RUN[EVENT == '3000S']),
      WIN_PCT_3000S = min(WIN_PCT[EVENT == '3000S']),
      AVG_TIME_FORM_3000S = min(AVG_TIME_FORM[EVENT == '3000S']),
      PR_FORM_3000S = min(PR_FORM[EVENT == '3000S']),
      # 5000m
      AVG_PLACE_5000 = min(AVG_PLACE[EVENT == '5000m']),
      AVG_TIME_5000 = min(AVG_TIME[EVENT == '5000m']),
      PR_5000 = min(PR[EVENT == '5000m']),
      WINS_5000 = min(WINS[EVENT == '5000m']),
      TIMES_RUN_5000 = min(TIMES.RUN[EVENT == '5000m']),
      WIN_PCT_5000 = min(WIN_PCT[EVENT == '5000m']),
      AVG_TIME_FORM_5000 = min(AVG_TIME_FORM[EVENT == '5000m']),
      PR_FORM_5000 = min(PR_FORM[EVENT == '5000m']),
      # 6K XC
      AVG_PLACE_6KXC = min(AVG_PLACE[EVENT == '6K XC']),
      AVG_TIME_6KXC = min(AVG_TIME[EVENT == '6K XC']),
      PR_6KXC = min(PR[EVENT == '6K XC']),
      WINS_6KXC = min(WINS[EVENT == '6K XC']),
      TIMES_RUN_6KXC = min(TIMES.RUN[EVENT == '6K XC']),
      WIN_PCT_6KXC = min(WIN_PCT[EVENT == '6K XC']),
      AVG_TIME_FORM_6KXC = min(AVG_TIME_FORM[EVENT == '6K XC']),
      PR_FORM_6KXC = min(PR_FORM[EVENT == '6K XC']),
      # 8K XC
      AVG_PLACE_8KXC = min(AVG_PLACE[EVENT == '8K XC']),
      AVG_TIME_8KXC = min(AVG_TIME[EVENT == '8K XC']),
      PR_8KXC = min(PR[EVENT == '8K XC']),
      WINS_8KXC = min(WINS[EVENT == '8K XC']),
      TIMES_RUN_8KXC = min(TIMES.RUN[EVENT == '8K XC']),
      WIN_PCT_8KXC = min(WIN_PCT[EVENT == '8K XC']),
      AVG_TIME_FORM_8KXC = min(AVG_TIME_FORM[EVENT == '8K XC']),
      PR_FORM_8KXC = min(PR_FORM[EVENT == '8K XC']),
      # 10K
      AVG_PLACE_10K = min(AVG_PLACE[EVENT == '10K']),
      AVG_TIME_10K = min(AVG_TIME[EVENT == '10K']),
      PR_10K = min(PR[EVENT == '10K']),
      WINS_10K = min(WINS[EVENT == '10K']),
      TIMES_RUN_10K = min(TIMES.RUN[EVENT == '10K']),
      WIN_PCT_10K = min(WIN_PCT[EVENT == '10K']),
      AVG_TIME_FORM_10K = min(AVG_TIME_FORM[EVENT == '10K']),
      PR_FORM_10K = min(PR_FORM[EVENT == '10K']),
    ) %>%
  mutate_if(is.numeric, list(~na_if(., Inf))) # Replace Inf values created by runners not competing in an event
  
  # Return data frame
  return(df)
}
```

# TIE DATA TO XC RESULTS
```{r}
# # Combine men and women from 2019
# xc19 = rbind(men_19, women_19)
# 
# # Reformat name in XC19
# xc19 = separate(xc19, NAME, c("LAST", "FIRST"), sep = ", ")
# xc19 = xc19 %>%
#   mutate(
#     NAME = toupper(paste0(FIRST, " ", LAST))
#   ) %>%
#   select(NAME, TEAM, YEAR, TIME, PL)
# 
# # Join runners info
# xc19 = xc19 %>%
#   left_join(runners, by = c("NAME"))
```

# Function to get data for URL
```{r}
getData = function(url){
  # URL to scrape
  meet_url = url
  # Get webpage URLs for runners
  meet_wp = getURLs(meet_url)
  # Query the runners results
  runners = resultsQuery(meet_wp)
  # Reformat the data
  runners = reformatRunners(runners)
  
  # Return final runners data frame
  return(runners)
}
```

# Get old XC data
```{r}
# URL to scrape
xc19 = 'https://www.tfrrs.org/results/xc/16731/NCAA_Division_I_Cross_Country_Championships'
runners_19 = getData(xc19)
# 2018
xc18 = 'https://www.tfrrs.org/results/xc/15036/NCAA_DI_Cross_Country_Championships'
runners_18 = getData(xc18)
# 2017
xc17 = 'https://www.tfrrs.org/results/xc/13423/NCAA_Division_I_Cross_Country_Championships'
runners_17 = getData(xc17)
# 2016
xc16 = 'https://www.tfrrs.org/results/xc/11271/NCAA_Division_I_Cross_Country_Championships'
runners_16 = getData(xc16)
# 2015
xc15 = 'https://www.tfrrs.org/results/xc/9347/NCAA_Division_I_Cross_Country_Championships'
runners_15 = getData(xc15)
# 2014
xc14 = 'https://www.tfrrs.org/results/xc/7652/NCAA_Division_I_Cross_Country_Championships'
runners_14 = getData(xc14)
# 2013
xc13 = 'https://www.tfrrs.org/results/xc/6218/NCAA_Division_I_Cross_Country_Championships'
runners_13 = getData(xc13)
# 2012
xc12 = 'https://www.tfrrs.org/results/xc/4797/2012_NCAA_Division_I_Cross_Country_Championships'
runners_12 = getData(xc12)
```

# Get D2 data
```{r}
# URL to scrape
xc19 = 'https://www.tfrrs.org/results/xc/16713/NCAA_Division_II_Cross_Country_Championships'
D2runners_19 = getData(xc19)
# 2018
xc18 = 'https://www.tfrrs.org/results/xc/15037/NCAA_Division_II_Cross_Country_Championships'
D2runners_18 = getData(xc18)
# 2017
xc17 = 'https://www.tfrrs.org/results/xc/13406/NCAA_Division_II_Cross_Country_Championships'
D2runners_17 = getData(xc17)
# 2016
xc16 = 'https://www.tfrrs.org/results/xc/11246/NCAA_Division_II_Cross_Country_Championships'
D2runners_16 = getData(xc16)
# 2015
xc15 = 'https://www.tfrrs.org/results/xc/9348/NCAA_Division_II_Cross_Country_Championships'
D2runners_15 = getData(xc15)
# 2014
xc14 = 'https://www.tfrrs.org/results/xc/7658/NCAA_Division_II_National_Championships'
D2runners_14 = getData(xc14)
# 2013
xc13 = 'https://www.tfrrs.org/results/xc/6221/NCAA_Division_II_Cross_Country_Championships'
D2runners_13 = getData(xc13)
# 2012
xc12 = 'https://www.tfrrs.org/results/xc/4796/2012_NCAA_Division_II_Cross_Country_Championships'
D2runners_12 = getData(xc12)
```


## Combine data & remove duplicated rows
### D1
```{r}
# Combine data
runners = rbind(runners_18, runners_19, runners_12, runners_13, runners_14, runners_15, runners_16, runners_17)

# Remove duplicates
runners = runners[!duplicated(runners), ]
```

### D2
```{r}
# Combine data
D2runners = rbind(D2runners_18, D2runners_19, D2runners_12, D2runners_13, D2runners_14, D2runners_15, D2runners_16, D2runners_17)

# Remove duplicates
D2runners = D2runners[!duplicated(D2runners), ]
```


## Write to CSV
```{r}
write.csv(runners, "Runner_Database.csv")
```






