---
title: "XC Modeling"
author: "Samuel Ivanecky"
date: "11/21/2019"
output: html_document
---
### SETUP
```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

### LOAD DATA
```{r}
setwd("~/Desktop/XC")
men = read.csv("MensXC.csv")
women = read.csv("WomensXC.csv")
```

### VARIABLE MANIPULATION
```{r}
# Create gender column
men$gender = "M"
women$gender = "F"
# Rename columns
names(women)[1] = "Place"
names(women)[2] = "School"
# Combine
df = rbind(men, women)
df = df[, c(1,2,5:14)]
# Convert everything to numeric
df$Score = as.numeric(df$Score)
df$X1 = as.numeric(df$X1)
df$X2 = as.numeric(df$X2)
df$X3 = as.numeric(df$X3)
df$X4 = as.numeric(df$X4)
df$X5 = as.numeric(df$X5)
# Subset data (remove 6-7 runners)
df = df[, c(1:8, 11, 12)]
```

### GROUP BY TEAM PLACE AND LOOK AT AVERAGE RUNNER PLACINGS
```{r}
avg_runners = df %>%
  group_by(Place) %>%
  summarise(R1 = round(mean(X1, na.rm = T)),
            R2 = round(mean(X2, na.rm = T)),
            R3 = round(mean(X3, na.rm = T)),
            R4 = round(mean(X4, na.rm = T)),
            R5 = round(mean(X5, na.rm = T)))

# Calculate spread
avg_runners$spread = avg_runners$R5 - avg_runners$R1
```

### WRTIE TO CSV
```{r}
#write.csv(avg_runners, "avg_runners.csv")
```

### SPLIT BACK INTO GENDERS
```{r}
men = df %>%
  filter(gender == 'M' & Place < 5)
women = df %>%
  filter(gender == 'F' & Place < 5)
```

### PODIUM FINISHES BY SCHOOL
```{r}
# Counts of schools by gender
ggplot(men, aes(School, fill = School)) + geom_bar(stat = "count") + ggtitle("Men")
ggplot(women, aes(School, fill = School)) + geom_bar(stat = "count") + ggtitle("Women")
```

### SUMMARY OF WHERE RUNNERS ON FIRST PLACE TEAMS FINISH
```{r}
# Subset to winning teams
winners = df %>%
  filter(Place == 1)
# Summaries of runners
summary(winners$X1)
summary(winners$X2)
summary(winners$X3)
summary(winners$X4)
summary(winners$X5)
```

# CALCULATE SPREAD
```{r}
df$spread = df$X5 - df$X1
```

### LOOK AT DISTRIBUTION OF SPREAD BY TEAM PLACE
```{r}
df$Place = as.factor(df$Place)
ggplot(df, aes(df$Place, df$spread)) + geom_boxplot() + ggtitle("Distribution of Spreads by Team Place") + labs(x = "Team Place", y = "Spread")
```

### CALCULATE GAPS BETWEEN RUNNERS
```{r}
# Get the gaps between each runner
df = df %>%
  mutate(
    g1 = X2 - X1, 
    g2 = X3 - X2, 
    g3 = X4 - X3,
    g4 = X5 - X4
  )
```

### GROUP TO AVG GAP LEVEL
```{r}
# Group down to the average level
avg_runners_gap = df %>%
  group_by(Place) %>%
  summarise(
    GAP1 = round(median(g1, na.rm = T)),
    GAP2 = round(median(g2, na.rm = T)),
    GAP3 = round(median(g3, na.rm = T)),
    GAP4 = round(median(g4, na.rm = T))
  )

# Convert to numeric
avg_runners_gap$Place = as.numeric(avg_runners_gap$Place)
```

### WRITE TO CSV
```{r}
write.csv(avg_runners_gap, "avg_gap.csv")
```

### CORRELATION TESTS
```{r}
# Attach data
attach(avg_runners_gap)

# Check correlations
cor.test(Place, GAP1)
cor.test(Place, GAP2)
cor.test(Place, GAP3)
cor.test(Place, GAP4)
```

### BREAK OUT TOP 10 TEAMS
```{r}
df$Place = as.numeric(df$Place)

# Subset data
top = df %>%
  filter(Place <= 8)

# Label podium teams
top = top %>%
  mutate(podium = 
           case_when(
             Place <= 4 ~ 'Y',
             T ~ 'N'
           ))
```

### LOOK AT AVGS OF PODIUM VS NON-PODIUM
```{r}
top_grp = top %>%
  group_by(podium) %>%
  summarise(
     r1 = round(median(X1)),
     r2 = round(median(X2)),
     r3 = round(median(X3)),
     r4 = round(median(X4)),
     r5 = round(median(X5)),
     g1 = round(median(g1)),
     g2 = round(median(g2)),
     g3 = round(median(g3)),
     g4 = round(median(g4)),
  )
```

```{r}
# Quick look at worst teams
worst = df %>%
  filter(Place == 31)

summary(worst$Score)
```

### AVG SCORES BY PLACE
```{r}
avg_scores = df %>%
  group_by(Place) %>%
  summarise(Avg_Score = mean(Score, na.rm = T), Med_Score = median(Score))
```

```{r}
write.csv(avg_scores, "avg_scores.csv")
```

```{r}
# Plot distributions of scores by team place
ggplot(df, aes(as.factor(Place), Score)) + geom_boxplot() + ggtitle("Distributions of Scores by Team Finish") + labs(x = "Team Finish", y = "Team Score")
```

###########################
### DRAFT RECAP CODE
###########################
### LOAD IN DATA
```{r}
ind = read.csv("Draft Results - Ind.csv")
team = read.csv("Draft Results - Team.csv")
```



